---
title: "Calculation of portfolios"
output: nils
---

```{r}
library(tidyquant)
library(tidyquant)
library(tidyverse)
library(dplyr)
library(plyr)
library(slider)
library(bsts)
library(matrixStats)
```

# ```{r}
#   library(doParallel)
#   library(snow)
#   library(doSNOW)
# cl <- makeCluster(detectCores()-1, type = "SOCK")
#  
# call_cluster_init = function(cl){
#  registerDoParallel(cl)
# }
# ```

```{r}
db <- dbConnect(RPostgres::Postgres(), dbname = "stock_db")

overview <- tbl(db, "overview")
daily = tbl(db, "daily")

df = daily %>% dplyr::select(stock_id, date, close) %>% as_tibble ()# connecting to db 

df$date = df$date %>% as.POSIXct()  

# df = overview %>% dplyr::filter(ipo_year == 2021) %>% dplyr::filter(exchange == "NASDAQ")
# df

# dbWriteTable(db, "daily", value = df, overwrite = TRUE, append= F, row.names = FALSE)

```


```{r}
##
## CREATING A MONTHLY DATASET ! 
##


#df_date = df_date as.Date.POSIXct(df_date, tryFormat = "%Y-%m-%d")

month_day = LastDayInMonth(df$date) %>% as.Date()

#elect (Convert(date, '2018-04-01'))
df$date = df$date %>% as.Date()  

df = df %>% filter(date == month_day) #Filters out all other dates for the month except last one

df$ranking_start = df$date %m-% months(12) # Start ranking of the portf, -12 months (as per MC paper)

day(df$ranking_start) = days_in_month(df$ranking_start) #Se day of data to last day of month

df$ranking_end = df$date %m-% months(2) # Per MC paper subtract two months

day(df$ranking_end) = days_in_month(df$ranking_end) # set day of the date = last day of the month

dfa= data.table::data.table(df) #making a "working" table
dfa$shifted_date<-dfa$date %m+% months(1) # Shifting the date with one month      
day(dfa$shifted_date)<-days_in_month(dfa$shifted_date) # Sets last day of months to shifted month  

dfa[is.na(dfa)] = 0 #set nas' to zero

month_day = NULL
```


```{r}
gaps= c(1) #c(1,7,30) # daily, weekly, monthly, return #### SET TO ONE AS THIS IS MONTHLY DATASET

dfa = dfa %>% group_by(stock_id) %>% arrange(date) #Ordering and sorting the data 

returns = function(gaps){(df$close - lead(df$close, n= gaps)) / lead(df$close,n = gaps)} # Return function, gives return per month pr stock (roling returns)

rets = as.data.frame(lapply(gaps,returns)) # Run function on the gaps(1) to get returns

#colnames(rets)[1] <- "d_ret" # Renaming cols 
#colnames(rets)[2] <- "w_ret"
colnames(rets)[1] <- "m_ret" # Gives correct names 

df_temp = dfa %>% # Adding cumalative returns 
  mutate(cr = cumprod(1 + rets$m_ret)) 

#dfa = cbind(dfa, d_ret = rets[1], w_ret = rets[2], m_ret = rets[3], cr = df_temp$cr) # binding all calculated data to dfa

dfa = cbind(dfa, d_ret = rets[1], cr = df_temp$cr) # Binding CR and monthly return to main dataset 

head(dfa) # Checking the DFA to make sure we got everything we need
```

```{r}
######
## OLD CODE, WAS USED TO CALC 1,7,30 DAYS RETURN. NOW SET IS MONTHLY.
###
# goal is to check for monthly 8 monthly returns over 11 month 

# gaps= c(1,7,30) # daily, weekly, monthly, return
# 
# dfa = dfa %>% group_by(symbol) %>% arrange(date) #Ordering and sorting the data 
# 
# returns = function(gaps){(df$close - lead(df$close, n= gaps)) / lead(df$close,n = gaps)} # Return function
# 
# rets = as.data.frame(lapply(gaps,returns))

#dfa$shifted_date<-dfa$date %m+% months(1) # Making a column with +1 month shift as of the time lag needed for formations date
```


```{r}
#names(dfa)[names(dfa) == "close"] = "close.x"

x = c("stock_id", "shifted_date", "m_ret") # creates variables to copy later  
return_cols = c() # Makes an empty vector where we get the returns 

#packs_needed = c("dplyr", "tidyverse") #redundant info, was built for foreach loop

for(i in c(2:12)){ # eleven months return 
    dfa$shifted_date = dfa$date %m+% months(i) # shift by i months
    day(dfa$shifted_date) = days_in_month(dfa$shifted_date) # Provides shift date with end of month
    
    df = merge(df, # merge return using the shifted date
                     dfa[x],
                     by.x=c("stock_id", "date"),
                     by.y=c("stock_id", "shifted_date"),
                     all.x=TRUE
                     )
    
    names(df)[names(df) == "m_ret.x"] <- "m_ret" # correct the column names
    names(df)[names(df) == "m_ret.y"] <-paste0("m_ret", i)# adding the roling returns for months  
    return_cols<-c(return_cols, paste0("m_ret", i)) # adds the new cols in a list to remove from set later
}

df # Checking dataset
return_cols # Checking return cols 
df$available_returns = rowSums(!is.na(df[,return_cols[2:11]])) # compute noumber of available returns
df

dfa=NULL # frees memory, dfa is no longer needed  
df$cum_return=rowProds(1+as.matrix(df[,return_cols[2:11]]), na.rm=T) # compute cumulative product
df = df[, -c(8:17)] # WARNING HARD CODED! Removes the m_rets previously calculated  

#df<-df[, !return_cols[2:11]] # drop return_cols (not needed anymore)
return_cols=NULL
df_temp = NULL
rets = NULL
month_day = NULL
glimpse(df) #checking the data
```


```{r}
library(data.table)
dfa = df

# remove rows which have available_returns < 8
dfa = dfa %>% filter(available_returns >= 8) # Provides the highest momentum stocks 
glimpse(dfa) # Rechecking table 

# normalize using the number of available returns
dfa$cum_return = dfa$cum_return / dfa$available_returns

# cut returns for each date into deciles to determine winner/loser portfolios
dfa$quantile = cut(dfa$cum_return,
    quantile(dfa$cum_return,probs=seq(from=0,to=1,by=1/10),na.rm =T),
    include.lowest=TRUE, 
    labels=FALSE)

top = dfa %>% filter(quantile == 10) # making the best performance portf 
top = top %>% arrange(date) # sorting after date  
top$cr_ret = top$close - lead(top$close) / lead(top$close) # Adds returns IDK if this will work here !!!!

bot = dfa %>% filter(quantile == 1) # "worst performing stocks 

#quant_portf = merge (bot, top, by.x = "date", by.y = "date") # just added them together (this is not used for anything yet)

#plotting the data, Get strange numbers. 
ggplot(top, aes(x = date)) +
       theme(legend.position = "top") +
       geom_smooth(aes(y = cum_return), color = "darkred") + 
       labs(x = 'Date',

```{r}
urlfile <- "https://raw.githubusercontent.com/emanuelaf/emanuelaf.github.io/master/data/sales-of-shampoo.csv"
shampoo_sales <- read_csv(urlfile)

```

```{r}
class(shampoo_sales$Month)
```

```{r}
shampoo_sales <- shampoo_sales %>%
  rename(time = Month, sales = `Sales of shampoo over a three year period`) %>%
  mutate(time = paste0("0", time)) %>%
  mutate(time = as.Date(parse_date_time(time, "%y%m"))) %>% 
  na.omit() 
#library(systemfonts)
install.packages("systemfonts")
knitr::kable(head(shampoo_sales), "html") %>%
  kableExtra::kable_styling(full_width = F, font_size = 11, position = "center")
```

```{r}
ggplot(shampoo_sales, aes(x = time, y = sales)) +
  geom_line(col = "hotpink") + ylim(0, 800) +
  scale_x_date(date_labels = "%y %b", date_breaks = "2 month") +
  theme_bw() + theme(legend.title = element_blank(),
                     axis.text.x  = element_text(angle=45, vjust=0.5))
```

```{r}
par(mfrow = c(1,2))
acf(as.ts(shampoo_sales$sales), main = "Sales")
pacf(as.ts(shampoo_sales$sales), main = "Sales")
```

```{r}
arima1 <- auto.arima(as.ts(shampoo_sales$sales))
arima1
```


##### 6 Month Prediction ####

```{r}
train_index <- 30
n_total <- nrow(shampoo_sales)
shampoo_sales_train1 <- shampoo_sales[1:(train_index),]
shampoo_sales_test <- shampoo_sales[(train_index+1):n_total,]
predicted <- numeric(n_total-train_index)
```

```{r}
for (i in 1:(n_total-train_index)) {
  shampoo_sales_train <- shampoo_sales[1:(train_index-1+i),]
  arima_model <- auto.arima(as.ts(shampoo_sales_train$sales))
  pred <- forecast(arima_model, 1)
  predicted[i] <- pred$mean
}
```

```{r}
df_pred <- tibble(obs = c(shampoo_sales_train1$sales, shampoo_sales_test$sales), 
                  predicted = c(shampoo_sales_train1$sales, predicted), 
                  time = shampoo_sales$time)
```

```{r}
ggplot(gather(df_pred, obs_pred, value, -time) %>% 
         mutate(obs_pred = factor(obs_pred, levels = c("predicted", "obs"))), 
       aes(x = time, y = value, col = obs_pred, linetype = obs_pred)) +
  geom_line() +
  xlab("") + ylab("") +
  scale_color_manual(values=c("black", "hotpink")) +
  scale_linetype_manual(values=c(2, 1)) +
  scale_x_date(date_labels = "%y %b", date_breaks = "1 month") +
  theme_bw() + theme(legend.title = element_blank(),
                     axis.text.x  = element_text(angle=45, vjust=0.5))
```
###### 12 Month Prediction ####

```{r}
train_index2 <- 24
n_total <- nrow(shampoo_sales)
shampoo_sales_train1 <- shampoo_sales[1:(train_index2),]
shampoo_sales_test <- shampoo_sales[(train_index2+1):n_total,]
predicted <- numeric(n_total-train_index2)
```


```{r}
for (i in 1:(n_total-train_index2)) {
  shampoo_sales_train <- shampoo_sales[1:(train_index2-1+i),]
  arima_model <- auto.arima(as.ts(shampoo_sales_train$sales))
  pred <- forecast(arima_model, 1)
  predicted[i] <- pred$mean
}
```

```{r}
df_pred <- tibble(obs = c(shampoo_sales_train1$sales, shampoo_sales_test$sales), 
                  predicted = c(shampoo_sales_train1$sales, predicted), 
                  time = shampoo_sales$time)
```

```{r}
ggplot(gather(df_pred, obs_pred, value, -time) %>% 
         mutate(obs_pred = factor(obs_pred, levels = c("predicted", "obs"))), 
       aes(x = time, y = value, col = obs_pred, linetype = obs_pred)) +
  geom_line() +
  xlab("") + ylab("") +
  scale_color_manual(values=c("black", "hotpink")) +
  scale_linetype_manual(values=c(2, 1)) +
  scale_x_date(date_labels = "%y %b", date_breaks = "1 month") +
  theme_bw() + theme(legend.title = element_blank(),
                     axis.text.x  = element_text(angle=45, vjust=0.5))
```

```{r}
Forecast <- forecast(arima_model, h=12)
plot(Forecast)
Forecast
```





       y = 'Cumulative Returns',
       title = 'Portfolio Cumulative Returns') +
    scale_y_continuous(breaks = seq(0,50000,5000)) +
    scale_x_date()#date_breaks = '10 year',
               #date_labels = '%Y')
##
##
## SOMEONE NEEDS TO CHECK THE FLOW ! `
##
##
```