---
title: "R Notebook"
output: html_notebook
---

dataloader

```{r}
library(dplyr)
library(doParallel)

```


```{r}
d.df = read_csv2("d_sp500.csv",col_names = TRUE)

tab_format = function (data_frame){
  test = data_frame %>% as_tibble()
  test = test %>% t()
  var1=NULL
  var2=NULL
  var3=NULL
  count = 1
  pb = txtProgressBar(min = 0, max = ncol(test), initial = 0, style = 3) # Progressbar 
    for (i in 1:ncol(test)) {
      var1 = test[,i] %>% as_tibble()
      test2 = test[1,i]
      var2 = var1 %>% mutate(symbol = test2)
      var3 = rbind(var2,var3)
      count = count +1
      vv = setTxtProgressBar(pb,count) # set the progressbar to update 
    }
  #var3 = slice(var3, n = 1)
  return(var3)
}

x = tab_format(d.df) 


```


Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Ctrl+Alt+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Ctrl+Shift+K* to preview the HTML file).

The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.
