---
title: "R Notebook"
output: html_notebook
---

```{r}
#install.packages("Quandl")
library(tidyquant)
library(tidyverse)
library(quantmod)
library(Quandl)
library(dplyr)
library(ggplot2)
library(tidyr)
library(zoo)
```


Getting dataset from https://mba.tuck.dartmouth.edu/pages/faculty/ken.french/data_library.html as described in MC paper.

```{r}

#Loading Kenneth french databases Average Value Weighted Returns -- Monthly
kf_monthly = read.csv("C:\\Users\\nilsp\\Dropbox\\Skole 2020 - 2021\\Machinelearning 1\\MomentumCrash\\momcrash\\10_Portfolios_Prior_12_2.CSV", header= TRUE, sep = ",")

#Loading Kenneth french databases Average Value Weighted Returns -- daily
kf_daily = read.csv("C:\\Users\\nilsp\\Dropbox\\Skole 2020 - 2021\\Machinelearning 1\\MomentumCrash\\momcrash\\10_Portfolios_Prior_12_2_daily.CSV", header= TRUE, sep = ",")
```


Dataset prepping 
```{r}
kf_monthly$X= paste(kf_monthly$X, "01", sep = '') # adding the first in a month to the date for easier processing
kf_monthly = transform(kf_monthly,X=as.Date(as.character(X),"%Y%m%d")) # converting X (dates) to date format
names(kf_monthly)[1]<-paste("Date") # changing name from X to Date to avoid confusion.
kf_monthly = kf_monthly %>% na.omit()

chars = sapply(kf_monthly, is.character) # Prepping to turn all data from char to num. 
kf_monthly[ , chars] = as.data.frame(apply(kf_monthly[ , chars], 2, as.numeric))

head(kf_monthly) # checking data
```


```{r}
  ggplot(kf_monthly, aes(x = Date)) +
    geom_line(aes(y = Lo.PRIOR), color = "darkred") + 
    geom_line(aes(y = PRIOR.2), color = "blue")
    labs(x = 'Date',
       y = 'Cumulative Returns',
       title = 'Portfolio Cumulative Returns') +
  theme_classic() +
  scale_y_continuous(breaks = seq(-1000,100,1000)) +
  scale_x_date(date_breaks = '10 year',
               date_labels = '%Y')
```


Prepping daily dataset
```{r}
kf_daily = transform(kf_daily,X=as.Date(as.character(X),"%Y%m%d")) # converting X (dates) to date format
names(kf_daily)[1]<-paste("Date") # changing name from X to Date to avoid confusion.
kf_daily = kf_daily %>% na.omit()

chars = sapply(kf_daily, is.character) # Prepping to turn all data from char to num. 
kf_daily[ , chars] = as.data.frame(apply(kf_daily[ , chars], 2, as.numeric))

head(kf_daily)
```
```{r}
  ggplot(kf_daily, aes(Date,Lo.PRIOR)) +
    geom_line(aes(y = Lo.PRIOR), color = "darkred") +
    geom_line(aes(y = PRIOR.2), color = "blue") +
    labs(x = 'Date',
       y = 'Cumulative Returns',
       title = 'Portfolio Cumulative Returns') +
  theme_classic() +
  scale_y_continuous(breaks = seq(-15,40,5)) +
  scale_x_date(date_breaks = '10 year',
               date_labels = '%Y')
```



# ```{r}
# Quandl.api_key("QnDmM5cUPLz3z6XAw5Uy") #Personal key -Nils
# Quandl('BSE/BOM504991')
# ```

<!-- # Import documentation for Quandl can be found here https://docs.data.nasdaq.com/ -->
<!-- ```{r} -->
<!-- mydata = Quandl("FRED/GDP", start_date="2001-12-31", end_date="2005-12-31") -->

<!-- mydataHK = Quandl("HKEX/00001") -->
<!-- ``` -->

No use for quandl as it is implemented in tidyquant package. 
Manual for tq can be found here 
https://cran.r-project.org/web/packages/tidyquant/vignettes/TQ01-core-functions-in-tidyquant.html

```{r}
#tickers = c('BND', 'VB', 'VEA', 'VOO', 'VWO', 'aapl', 'msft')

#Old stocks pre year 1911
tickers = c('xom', 'ge', 'cvx', 'ibm', 'cl', 'pg', 'jnj', 'jpm','pfe', 'ko', 'wfc', 'c', 'bac', 'mrk', 'pep', 'abt', 'gs', 'ups', 'mmm', 'axp', 'f', 'cvs', 'usb', 'unp', 'bmy', 'dd', 'met', 'lly', 'dow', 'EMR')

# Asset weights

wts = c(0.033,0.033,0.033,0.033,0.033,0.033,0.033,0.033,0.033,0.033,0.033,0.033,0.033,0.033,0.033,0.033,0.033,0.033,0.033,0.033,0.033,0.033,0.033,0.033,0.033,0.033,0.033,0.033,0.033,0.033) #All weights sett to 1/30, due to copy paste

price_data <- tq_get(tickers,
                     from = '1969-01-01',
                     to = '2020-12-31',
                     get = 'stock.prices')

ret_data <- price_data %>%
  group_by(symbol) %>%
  tq_transmute(select = adjusted, #NB! always use adjusted! Due to close does not take into considiration stocksplits dividents etc. 
               mutate_fun = periodReturn,
               period = "daily",
               col_rename = "ret")

wts_tbl <- tibble(symbol = tickers,
                  wts = wts)
ret_data <- left_join(ret_data,wts_tbl, by = 'symbol')

ret_data <- ret_data %>%
  mutate(wt_return = wts * ret)

port_ret <- ret_data %>%
  group_by(date) %>%
  summarise(port_ret = sum(wt_return))

port_cumulative_ret <- port_ret %>%
  mutate(cr = cumprod(1 + port_ret))
```

```{r}
# same as above just shorter ad prettier
RLS = NULL
RSS = NULL
RFT = NULL


port_ret_tidyquant <- ret_data %>%
  tq_portfolio(assets_col = symbol,
               returns_col = ret,
               weights = wts,
               col_rename = 'port_ret',
               geometric = FALSE)
# Cumalative return calculation p.23 eq. 12-14 in MC paper 
port_cumulative_ret_tidyquant <- port_ret_tidyquant %>%
  mutate(cr = cumprod(1 + port_ret)-1) %>%
  mutate(logret = cumsum(log(1+port_ret)))  #%>%
  #mutate(longshort = cumprod(1+RLS+RSS+RFT)-1)  # RLS = return long side of pf RSS = ret short side RFT = return risk free rate 
```

```{r}
port_cumulative_ret %>%
  mutate(port_ret_tidyquant = port_cumulative_ret_tidyquant$cr) %>%
  select(-port_ret) %>%
  rename(long_method = cr) %>%
  gather(long_method,port_ret_tidyquant,
         key = port_method,
         value = cr) %>%
  ggplot(aes(x = date, y = cr, color = port_method)) +
  geom_line() +
  labs(x = 'Date',
       y = 'Cumulative Returns',
       title = 'Portfolio Cumulative Returns') +
  theme_classic() +
  scale_y_continuous(breaks = seq(0,600,25)) +
  scale_x_date(date_breaks = 'year',
               date_labels = '%Y')
```

```{r}
#lag(adjClose) / lag(adjClose, 12) - 1)
#formula for adjusted return
```


```{r}
#Return.cumulative(port_cumulative_ret$cr, geometric = FALSE) #OBJECT placeholder is data to calculate!
```

Docs can be found at https://www.rdocumentation.org/packages/PerformanceAnalytics/versions/2.0.4/topics/Return.cumulative. For this exercise this is equal to the equation p.23 eq. 12. Geometric is set to false to provide arithmatic returns 

More on cumalative return calculations can be found here https://www.codingfinance.com/post/2018-04-10-cumulative-portfolio-returns/
```{r}

```


Sharp ratio https://cran.r-project.org/web/packages/SharpeR/vignettes/SharpeR.pdf